---
title: "Spectral Signatures and Vegetation Indices (Sentinel-2)"
author: "Tom Ihro"
date: "2023-11-14"
output: html_document
---

```{r}

library(terra)

img = rast("data/sen2_20200821_sen2b.tif")
img = img/100 #setting for reflectivity
setMinMax(img)

#set negative values to nan
#img[img < 0] = NaN

plotRGB(img,4,3,2,stretch = "lin")
```

```{r warning=FALSE,}
landcover <- vect("data/landcoverNew.kml")
landcover = project(landcover, img)

writeVector(landcover,"data/landcoverclasses.gpkg",overwrite = TRUE)


plotRGB(img, r=8, g=3, b=2, stretch="lin")
plot(landcover, add=TRUE, col = "green")
```

```{r,}

img.extract <- extract(img, landcover, fun=mean)

names(img.extract) <- c("ID", "blue", "green", "red", "RE1", "RE2", "RE3", "NIR_b", "NIR_a", "SWIR1", "SWIR2")
row.names(img.extract) <- c(landcover$Name)

img.extract
```

```{r}
img.extract[1,2:11]
```

```{r}

lambda<-c(490,560, 665, 705, 740, 783, 842, 865, 1610, 2190)

landcover_col <- c("forest green", 
                   "blue1", 
                   "light blue", 
                   "grey", 
                   "light green",
                   "beige",
                   "steelblue",
                   "red",
                   "darkseagreen",
                   "green3")

plot(lambda,img.extract[1,2:11],
     col = "forest green",
     xlab="Wavelength [nm]", 
     ylab="reflectance (%)", 
     main="reflectance spectra",
     lwd=2, type='b',
     xlim=c(400,2700), 
     ylim=c(0,40))

#loop through all landcover classes
for(i in 2:10){
  lines(lambda, img.extract[i,2:11], col = landcover_col[i],lwd=2, type='b')
}

legend('topright', title='Landcover', 
       legend=row.names(img.extract),
       fill=landcover_col,
       cex = 0.8, bty="n")
```

```{r, warning=F}

plot(img[[2]], img[[3]],  xlab="green", ylab="red", xlim=c(0,40), ylim=c(0,40), maxcell = 100000,  col= "gray35")

ran = 1:5

for(i in ran){
  points(img.extract[i,"green"],img.extract[i,"red"], col=landcover_col[i], cex=2, pch=19)
}

legend('topright', title='Landcover', 
       legend=row.names(img.extract)[ran],
       fill=landcover_col[ran],
       cex = 0.8, bty="n") 

```

# Homework

## 1. Feature Space

```{r, warning=FALSE}
plot(img[[3]], img[[8]],  xlab="red", ylab="NIR", xlim=c(0,40), ylim=c(0,40), maxcell = 100000,  col= "gray35")

ran = 1:5 #Define the range of which classes we are looking at

for(i in ran){
  points(img.extract[i,"red"],img.extract[i,"NIR_a"], col=landcover_col[i], cex=2, pch=19)
}

legend('topright', title='Landcover', 
       legend=row.names(img.extract)[ran],
       fill=landcover_col[ran],
       cex = 0.8, bty="n") 
```

```{r, warning=FALSE}

plot(img[[8]], img[[9]],  xlab="NIR_a", ylab="SWIR1", xlim=c(0,40), ylim=c(0,40), maxcell = 100000,  col= "gray35")

ran = 1:5 #Define the range of which classes we are looking at

for(i in ran){
  points(img.extract[i,"NIR_a"],img.extract[i,"SWIR1"], col=landcover_col[i], cex=2, pch=19)
}

legend('topright', title='Landcover', 
       legend=row.names(img.extract)[ran],
       fill=landcover_col[ran],
       cex = 0.8, bty="n") 
```

Unlike for the Visual bands for the NIR and SWIR bands the bands seem to not be correlated that much. As one can observe by the differing Feature spaces the higher

## NDVI

```{r}

NDVI = (img[[8]] - img[[3]])/(img[[8]] + img[[3]])

#adopt range -1,1 as that is the range that the NDVI function works on. 
plot(NDVI,stretch = "lin",range = c(-1,1))
```

```{r}

NDVI.class = NDVI

#Classify to 1 -> Water 2 -> Land
NDVI.class[NDVI < 0] = 1
NDVI.class[NDVI >= 0] = 2

color = data.frame(value = c(1,2), color = c("lightblue","brown"))

plot(NDVI.class,col = color$color,type = "classes",levels = c("Water","Land"))
```

## EVI

```{r}

G = 2.5
C1 = 6
C2 = 7.5
L = 1

rhoNIR = img$sen2_20200821_sen2b_7/100
rhoRED = img$sen2_20200821_sen2b_3/100
rhoBLUE = img$sen2_20200821_sen2b_1/100


EVI = G*((rhoNIR - rhoRED)/(rhoNIR + C1*rhoRED - C2*rhoBLUE+L))


plot(EVI,range = c(-0.2,1))

```

## PVI

```{r, warning=FALSE}

X = c(5,25)
Y = c(8,23)


par(mfrow = c(1,1), pty = "s")
plot(img[[3]], img[[8]],  xlab="red", ylab="NIR", xlim=c(0,40), ylim=c(0,40), maxcell = 100000,  col= "gray35")
grid()
points(X,Y, col = "red")
```

```{r}

a = (Y[2] - Y[1])/(X[2]-X[1])
b = Y[1] - a*X[1]


PVI = (img[[8]] - a*img[[3]]-b)/sqrt(1+a^2)

plot(PVI,range=c(-10,40))
```

## Compare Indexes

```{r}


```

```{r, warning=FALSE}


plot(NDVI,EVI,maxcell = 100000,
     xlim=c(-0.2,1),
     ylim=c(-1,1),
     xlab ="NDVI",
     ylab ="EVI",
     )

```

```{r, warning=FALSE}
plot(NDVI,PVI,maxcell = 1000000,
     xlim=c(-1,1),
     ylim=c(-10,40),
     xlab ="NDVI",
     ylab ="PVI",
     )
```

```{r, warning=FALSE}
plot(EVI,PVI,maxcell = 100000,
     xlim=c(-1,1),
     ylim=c(-10,40),
     xlab ="EVI",
     ylab ="PVI",
     )
```

We can see that especially with the EVI vs PVI scatter plot that these two seem to correlate the most. Compared to the NDVI vs PVI which has alot of degeneracy at $PVI \approx -4$ which means that there can be alot of NDVI values corresponding to that specific PVI value.
